---
title: "STA380"
author: "SebastianOsorio"
date: "12/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(mosaic)
library(quantmod)
library(foreach)
library(xts)
```

	
# Import ETFs

SPDR S&P 500 (SPY): The oldest surviving and most widely known ETF tracks the S&P 500 Index
iShares Russell 2000 (IWM): Tracks the Russell 2000 small-cap index
Invesco QQQ (QQQ): Indexes the Nasdaq 100, which typically contains technology stocks
SPDR Dow Jones Industrial Average (DIA): Represents the 30 stocks of the Dow Jones Industrial Average
Sector ETFs: Track individual industries such as oil (OIH), energy (XLE), financial services (XLF), REITs (IYR), Biotech (BBH)
Commodity ETFs: Represent commodity markets including crude oil (USO) and natural gas (UNG)
Physically-Backed ETFs: The SPDR Gold Shares (GLD) and iShares Silver Trust (SLV) hold physical gold and silver bullion in the fund

```{r}
#First Portfolio : "Tech"
	ETFs = c("QQQ","DIA","USO","SPY", "GLD")
	getSymbols(ETFs, from = "2014-01-01")
```

	# Adjusting for splits and/or dividends

```{r}
	QQQa = adjustOHLC(QQQ)
	DIAa = adjustOHLC(DIA)
	USOa = adjustOHLC(USO)
	SPYa = adjustOHLC(SPY)
	GLDa = adjustOHLC(GLD)
```
	
	# Combine close to close changes in a single matrix

```{r}
	all_etf = cbind(ClCl(QQQa),ClCl(DIAa),ClCl(USOa),ClCl(SPYa),ClCl(GLDa))
```


```{r}
	all_etf = as.matrix(na.omit(all_etf))
```

```{r}
for(ticker in ETFs) {
	expr = paste0(ticker, "a = adjustOHLC(", ticker, ")")
	eval(parse(text=expr))
}
```


```{r}
# Combine all the returns in a matrix
all_etf = cbind(	ClCl(QQQa),
								ClCl(DIAa),
								ClCl(USOa),
								ClCl(SPYa),
								ClCl(GLDa))
head(all_etf)
all_etf = as.matrix(na.omit(all_etf))
```

```{r}

# Now simulate many different possible scenarios  
initial_wealth = 100000
sim1 = foreach(i=1:5000, .combine='rbind') %do% {
	total_wealth = initial_wealth
	weights = c(0.2, 0.2, 0.2, 0.2, 0.2)
	holdings = weights * total_wealth
	n_days = 20
	wealthtracker = rep(0, n_days)
	for(today in 1:n_days) {
		return.today = resample(all_etf, 1, orig.ids=FALSE)
		holdings = holdings + holdings*return.today
		total_wealth = sum(holdings)
		wealthtracker[today] = total_wealth
	}
	wealthtracker
}

# Profit/loss and VAR
Port1_mean = mean(sim1[,n_days])
Port1_VAR = quantile(sim1[,n_days],.05)
Port1_Upper = quantile(sim1[,n_days],.95)
```


```{r Second Portfolio : Commodities and phisically backed}
#Second Portfolio : "Commodities and phisically backed"
	ETFs = c("GLD","SLV","USO", "UNG")
	getSymbols(ETFs, from = "2014-01-01")
```

	# Adjusting for splits and/or dividends

```{r}
	GLDa = adjustOHLC(GLD)
	SLVa = adjustOHLC(SLV)
	USOa = adjustOHLC(USO)
	UNGa = adjustOHLC(UNG)

```
	

	# Combine close to close changes in a single matrix

```{r}
	all_etf = cbind(ClCl(GLDa),ClCl(SLVa),ClCl(USOa),ClCl(UNGa))
```


```{r}
	all_etf = as.matrix(na.omit(all_etf))
	head(all_etf)
	pairs(all_etf)
	dim(all_etf)
```

```{r}
for(ticker in ETFs) {
	expr = paste0(ticker, "a = adjustOHLC(", ticker, ")")
	eval(parse(text=expr))
}

head(GLDa)
```


```{r}
# Combine all the returns in a matrix
all_etf = cbind(	ClCl(GLDa),
								ClCl(SLVa),
								ClCl(USOa),
								ClCl(UNGa))
head(all_etf)
all_etf = as.matrix(na.omit(all_etf))
```


```{r}
# Now simulate many different possible scenarios  
initial_wealth = 100000
sim2 = foreach(i=1:5000, .combine='rbind') %do% {
	total_wealth = initial_wealth
	weights = c(0.25, 0.25, 0.25, 0.25)
	holdings = weights * total_wealth
	n_days = 20
	wealthtracker = rep(0, n_days)
	for(today in 1:n_days) {
		return.today = resample(all_etf, 1, orig.ids=FALSE)
		holdings = holdings + holdings*return.today
		total_wealth = sum(holdings)
		wealthtracker[today] = total_wealth
	}
	wealthtracker
}

# Profit/loss
Port2_Mean = mean(sim2[,n_days])
Port2_VAR = quantile(sim2[,n_days],.05)
Port2_Upper = quantile(sim2[,n_days],.95)
```


```{r}
#p of positive returns
mean(sim1[,n_days]>100000)
```

SPDR S&P 500 (IYR): The oldest surviving and most widely known ETF tracks the S&P 500 Index
iShares Russell 2000 (IWM): Tracks the Russell 2000 small-cap index
Invesco OIH (OIH): Indexes the Nasdaq 100, which typically contains technology stocks
SPDR Dow Jones Industrial Average (XLE): Represents the 30 stocks of the Dow Jones Industrial Average
Sector ETFs: Track individual industries such as oil (OIH), energy (XLE), financial services (XLF), REITs (IYR), Biotech (BBH)
Commodity ETFs: Represent commodity markets including crude oil (XLF) and natural gas (UNG)
Physically-Backed ETFs: The SPDR Gold Shares (BBH) and iShares Silver Trust (SLV) hold physical gold and silver bullion in the fund

```{r}
#Third Portfolio : "Sector ETFs"
	ETFs = c("OIH","XLE","XLF","IYR", "BBH")
	getSymbols(ETFs, from = "2014-01-01")
```


	

	# Adjusting for splits and/or dividends

```{r}
	OIHa = adjustOHLC(OIH)
	XLEa = adjustOHLC(XLE)
	XLFa = adjustOHLC(XLF)
	IYRa = adjustOHLC(IYR)
	BBHa = adjustOHLC(BBH)
```
	

	# Combine close to close changes in a single matrix

```{r}
	all_etf = cbind(ClCl(OIHa),ClCl(XLEa),ClCl(XLFa),ClCl(IYRa),ClCl(BBHa))
```


```{r}
	all_etf = as.matrix(na.omit(all_etf))
	head(all_etf)
	pairs(all_etf)
	dim(all_etf)
```

```{r}
for(ticker in ETFs) {
	expr = paste0(ticker, "a = adjustOHLC(", ticker, ")")
	eval(parse(text=expr))
}

head(OIHa)
```


```{r}
# Combine all the returns in a matrix
all_etf = cbind(	ClCl(OIHa),
								ClCl(XLEa),
								ClCl(XLFa),
								ClCl(IYRa),
								ClCl(BBHa))
head(all_etf)
all_etf = as.matrix(na.omit(all_etf))
```


```{r}

# Now simulate many different possible scenarios  
initial_wealth = 100000
sim3 = foreach(i=1:5000, .combine='rbind') %do% {
	total_wealth = initial_wealth
	weights = c(0.2, 0.2, 0.2, 0.2, 0.2)
	holdings = weights * total_wealth
	n_days = 20
	wealthtracker = rep(0, n_days)
	for(today in 1:n_days) {
		return.today = resample(all_etf, 1, orig.ids=FALSE)
		holdings = holdings + holdings*return.today
		total_wealth = sum(holdings)
		wealthtracker[today] = total_wealth
	}
	wealthtracker
}

# Profit/loss and VAR
Port3_Mean = mean(sim3[,n_days])
Port3_VAR = quantile(sim3[,n_days],.05)
Port3_Upper = quantile(sim3[,n_days],.95)

```


```{r}
#Compare portfolios:
par(mfrow = c(3, 1))
hist(sim1[,n_days]- initial_wealth, breaks=30, xlim = range(-30000, 30000), ylim = range(0, 1000), main = "P1: Tech", xlab = "$ Total Returns")
hist(sim2[,n_days]- initial_wealth, breaks=30, xlim = range(-30000, 30000), ylim = range(0, 1000), main = "P2: Commodities and Phisically Backed", xlab = "$ Total Returns")
hist(sim3[,n_days]- initial_wealth, breaks=30, xlim = range(-30000, 30000), ylim = range(0, 1000), main = "P3: Sector ETFs", xlab = "$ Total Returns")

```

```{r}
#Compare average estimated returns
Port1_Mean
Port2_Mean
Port3_Mean

#Compare the VAR
Port1_VAR
Port2_VAR
Port3_VAR

#Compare Upper Returns
Port1_Upper
Port2_Upper
Port3_Upper
```
Comparing the three portfolios of P1 Tech, P2 Commodities and Phisically Backed (C&PB) and P3 Sector ETFs shows similar distributions. Inspecting the histograms, Tech appears to have a normal distribution, C&PB has a left tail and Sector ETFs has a right tail.

Tech has the highest average return of 546.80 dollars above the initial 100,000 dollars. Sector ETFs also had a positive average return of 368.5 dollars while C&PB had an average negative return of 822.94 dollars. 

The VAR for each portfolio was 95,052.90, 91,488.25 and 93,114.5 dollars for Tech, Sector ETFS and C&PB respectivley. Given the left tail for C&PB it is unsurprising that it has the lowest VAR. Tech has the best VAR - which would be the preferable portfolio for a risk averse person. 

Conversley, comparing the upper 95% simulated quintile shows C&PB has the highest upside with a return of 8,705.30 dollars, second to Sector EFTs with 8,443.70 dollars. Tech has the lowest return at the 95% quintile of 6,246.00 dollars.

Given Tech has the highest VAR and average return, a risk averse person should invest in this portfolio. For a risk seeker, they may try the C&PB portfolio which has the highest variance and highest potential upside.