---
title: "Market Segmentation"
author: "Rocco Lange"
date: "8/15/2019"
output: 
  md_document:
    variant: gfm
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# MARKET SEGMENTATION:  
## Analysis:  

In order to better understand NutrientH20’s Twitter followers, Principal Component Analysis was used. Before running the algorithm, the Tweets under the ‘chatter’ and ‘uncategorized’ categories were removed as they do not provide any useful information for a market segmentation. Next, the data was normalized to use the percentage of Tweets not the overall number of Tweets in each category.  PCA was run to obtain 3 principal components, as these explain 35% of the variance in the data, to understand the categories these people tweeted about and assign them to a client archetype.  
 
 
## Findings:  

| Component | Most Tweeted                                               |
|-----------|------------------------------------------------------------|
| PC1       | Health and Nutrition, Personal Fitness, Cooking, Outdoors  |
| PC2       | Photo Sharing, Cooking, Shopping, Fashion                  |
| PC3       | College/University, Online Gaming, Sports Playing, TV/Film |


According to these components, a couple archetypes emerged.  

**1. The outdoorsy:** This segment corresponds to the followers that are high on component 1. They tend to be health conscious people that are concerned with their fitness, which means they are a great target for NutrientH20 for its conceived purpose of hydration.  

**2. The social:** Follows trends and want to share it with their network. They are buying the brand because it is perceived as cool or new. They probably want to show the world they try new things.  

**3. The college student:** They drink NutrientH20 just because it is close by or because of the taste. They don't really associate the brand to a nutricious or sports related drink.  

```{r, include = FALSE}
rm(list = ls())
market = read.csv("social_marketing.csv", row.names=1)

library(tidyr)
library(mosaic)
library(dplyr)
library(reshape2)
library(ggplot2)
```

```{r, echo = FALSE}
market <- cbind(market[,2:4],market[6:36])
market_scale = market
cormat <- round(cor(market), 2)
melted_cormat <- melt(cormat)

#market_scale = sweep(market, MARGIN = 1, FUN = '/', STATS = rowSums(market[,]))
market_scale = market/rowSums(market)
market_scale = scale(market_scale, center = TRUE, scale = FALSE)

pc3 = prcomp(market_scale, scale=FALSE, rank=3)
#plot(pc5)
loadings = pc3$rotation
scores = pc3$x
#head(market_scale)
pairs(scores)

```

	
## Conclusion:  
Using this analysis, we believe that these three groups represent the market segments for NutrientH2O. Each group should be marketed to in a different way. The first group cares about their health and sports performance, they will buy NutrientH20 by associating the brand to high performance sports figures.  The second segment, is harder to market to as they will drink it if enough people do it or if it is perceived as cool. It is possible to use fashion figures or artists. A good way to accomodate for both would be to get a sports and fashion forward figure such as David Beckham as a brand ambassador. The last market segment is all about being a "healthy" alternative to sugar sodas. They will buy it if it is readily available, and getting NutrientH2O placed in vending machines and/or sold at markets around high schools and college campuses.

Lastly, it is important to remark that all three segments have politics and travel among their least tweeted topics. A good recommendation would be to stay away from politics as these followers are not concerned with it and could possibly be alienated by politics. On the other hand, trying to associate the brand with travelling might bring forward new segments.


